Nodal Discontinuous Galerkin Method for the Two-Moment Model of Neutrino Transport

### 2017-07-22 Documents/M1/dgM1Curvilinear.tex
Minor updates.

As we mentioned in Section~\ref{sec:twoMomentModel} when discussing Eqs.~\eqref{eq:momentEquationEnergy} and \eqref{eq:momentEquationMomentum}, the total opacity (excluding NES) is $\kappa(\epsilonNu,\bU)=\tilde{\chi}(\epsilonNu,\bU)+\sigma(\epsilonNu,\bU)$, where the absorption opacity $\tilde{\chi}(\epsilonNu,\bU)$, corrected for stimulated absorption, is given by the sum of the neutrino emissivity $\eta(\epsilonNu,\bU)$ and the neutrino absorptivity $\chi(\epsilonNu,\bU)$
\begin{equation}
  \tilde{\chi}(\epsilonNu,\bU) = \eta(\epsilonNu,\bU) + \chi(\epsilonNu,\bU). 
\end{equation}
The scattering opacity $\sigma(\epsilonNu,\bU)$ is due to (isotropic) elastic scattering.   
Neutrino-electron scattering gives rise to the integral operators involving the rates $R^{\In}(\epsilonNu,\epsilonNu',\bU)$ and $R^{\In}(\epsilonNu,\epsilonNu',\bU)$.  
(In the following, to simplify the notation, we will drop the explicit spatial dependence and dependence on material properties $\bU$.)  
To arrive at the collision terms for the two-moment model, we consider the space-homogeneous Boltzmann equation \citep[e.g., Eq.~(7) in][]{mezzacappaBruenn_1993c}
\begin{align}
  \f{1}{c}\deriv{f}{t}(\omegaNu,\epsilonNu,t)
  & = \eta(\epsilonNu) - \tilde{\chi}(\epsilonNu)\,f(\omegaNu,\epsilonNu,t) \nonumber \\
  & \hspace{12pt}
  + \frac{\epsilonNu^2}{c\,(hc)^3} \int_{\bbS^{2}} R_{\IS}(\bl'\cdot\bl,\epsilonNu)\,f(\omega', \epsilonNu,t)\,d\omegaNu'
  - \frac{\epsilonNu^2}{c\,(hc)^3} f(\omegaNu,\epsilonNu,t) \int_{\bbS^{2}} R_{\IS}(\bl'\cdot\bl,\epsilonNu)\,d\omega' \nonumber \\
  & \hspace{12pt} 
  + \frac{1}{c\,(hc)^3}\big(1- f(\omegaNu,\epsilonNu)\big) \, 
  \int_{\bbR^{+}} \int_{\bbS^{2}}  R^{\In}_{\NES} (\bl'\cdot\bl,\epsilonNu,\epsilonNu')\,f(\omega',\epsilonNu',t)\,d\epsilonNu' \epsilonNu'^{2}\,d\omega' \nonumber \\
  & \hspace{12pt} 
  - \frac{1}{c\,(hc)^3} f(\omegaNu,\epsilonNu,t) \int_{\bbR^{+}} 
  \int_{\bbS^{2}}  R^{\Out}_{\NES} (\bl'\cdot\bl,\epsilonNu,\epsilonNu^{\prime})\,\big(1-f(\omegaNu',\epsilonNu',t )\big)\,d\epsilonNu' \epsilonNu'^{2}\,d\omegaNu',
  \label{eq:homogeneousBoltzmann}
\end{align}
where we have brought back physical constants; i.e., the speed of light $c$ and the Planck constant $h$.  
In the scattering kernels, $\bl(\omegaNu)=(\cos\vartheta,\sin\vartheta\cos\varphi,\sin\vartheta\sin\varphi)^{\mbox{\tiny T}}$, so that
\begin{equation}
 \bl'\cdot\bl  = \mu\mu' + [(1-\mu^2)(1-\mu'^{2})]^{1/2}\cos(\phiNu-\phiNu')
 \equiv \cos\alpha,
 \quad(\mu=\cos\thetaNu)
\end{equation}
is the cosine of the angle between incoming and outgoing neutrino is the scattering process.  
In Eq.~\eqref{eq:homogeneousBoltzmann}, the first two terms are due to emission and absorption, the third and fourth terms are due to elastic (isoenergetic) neutrino scattering on nucleons and nuclei, and the last two terms are due to inelastic (non-isoenergetic) neutrino-electron scattering.  
In the following subsections we separately give the collision terms for the two-moment model by taking angular moments of Eq.~\eqref{eq:homogeneousBoltzmann}.  
The total collision term in the two-moment model is then given by the sum of the separate terms.

@@

Electron capture on nucleons and nuclei ($p+e^{-}\rightleftharpoons n+\nu_{e}$) is governed by the corrected absorption opacity $\tilde{\chi}(\epsilonNu)$ 
\begin{equation}
  \tilde{\chi}(\epsilonNu) = \chi(\epsilonNu) + \eta(\epsilonNu), 
\end{equation}
where $\eta(\epsilonNu)$ is the emissivity (taken as the sum of contributions from nucleons and nuclei)
\begin{equation}
  \eta = \eta_{\mbox{\tiny nucleon}} + \eta_{\mbox{\tiny nuclei}},
\end{equation}
and $ \chi(\epsilonNu)$ is the absorption opacity.  
They are related by equilibrium condition
\begin{equation}
  \chi(\epsilonNu) = e^{(\epsilonNu-\mu_{\nu})/k_{\mbox{\tiny B}}T}\,\eta(\epsilonNu). 
\end{equation}
Therefore,
\begin{equation}
  \chi = e^{(\epsilonNu-\mu_{\nu})/k_{\mbox{\tiny B}}T}\,(\eta_{\mbox{\tiny nucleon}} + \eta_{\mbox{\tiny nuclei}}) 
  \equiv \chi_{\mbox{\tiny nucleon}} + \chi_{\mbox{\tiny nuclei}},
\end{equation}
and
\begin{equation}
  \eta-\tilde{\chi}\,f=\tilde{\chi}\,\big(\,\eta/\tilde{\chi}-f\,\big)=\tilde{\chi}\,\big(\,f_{0}-f\,\big).  
\end{equation}
Thus, the corresponding moment equations are
\begin{equation}
  \f{1}{c}\deriv{\cJ}{t}=\tilde{\chi}\,\big(\,\cJ_{0}-\cJ\,\big)
  \quad\text{and}\quad
  \f{1}{c}\deriv{\bcH}{t}=-\tilde{\chi}\,\bcH.  
\end{equation}
We use the electron-neutrino nucleon absorption opacity $\chi_{\mbox{\tiny nucleon}}$ and emissivity $\eta_{\mbox{\tiny nucleon}}$ as given by Eq.~(C13) and Eq.~(C15), respectively, in \citet{bruenn_1985}, and the electron-neutrino nuclei absorption opacity $\chi_{\mbox{\tiny nuclei}}$ and emissivity $\eta_{\mbox{\tiny nuclei}}$ as given by Eq.~(C29) and Eq.~(C27), respectively, in \citet{bruenn_1985}.  
We only tabulate the corrected absorption opacity $\tilde{\chi}$, which has units cm$^{-1}$.

@@

The isoenergetic scattering kernel $R_{\IS}(\bl'\cdot\bl,\epsilonNu)$ is taken as the sum of contributions from scattering on nucleons and nuclei
\begin{equation}
  R_{\IS}(\bl'\cdot\bl,\epsilonNu,\epsilonNu) = R_{\IS}^{\mbox{\tiny nucleon}}(\bl'\cdot\bl,\epsilonNu) + R_{\IS}^{\mbox{\tiny nuclei}}(\bl'\cdot\bl,\epsilonNu),
\end{equation}
where $R_{\IS}^{\mbox{\tiny nucleon}}$ and $R_{\IS}^{\mbox{\tiny nuclei}}$ are given by Eq. (35) and Eq. (36), respectively, in \citet{mezzacappaBruenn_1993a}.  
For inclusion in the two-moment model, these kernels are approximated by a two-term Legendre series
\begin{align}
  R_{\IS}(\cos\alpha,\epsilonNu) \approx \f{1}{2} \Phi_{\IS,0}(\epsilonNu) + \f{3}{2}\Phi_{\IS,1}(\epsilonNu)\,\cos\alpha.
\end{align}
By orthogonality of the Legendre polynomials ($P_{0}(x)=1,P_{1}(x)=x$), the expansion coefficients are given by
\begin{equation}
  \Phi_{\IS,\ell}(\epsilonNu)
  =\int_{-1}^{+1}R_{\IS}(\cos\alpha,\epsilonNu)\,P_{\ell}(\cos\alpha)\,d\cos\alpha \quad (\ell=0,1).  
\end{equation}
Inserting the Legendre expansion into Eq.~\eqref{eq:homogeneousBoltzmann} and taking the angular moments gives no contribution to the equation for $\cJ$ (energy conservation), while the contribution to the equation for $\bcH$ is
\begin{equation}
  \f{1}{c}\deriv{\bcH}{t}
  =-\big(\,\sigma_{\IS,0}-\sigma_{\IS,1}\big)\,\bcH,
\end{equation}
where we have defined
\begin{equation}
  \sigma_{\IS,\ell}(\epsilonNu)
  =\f{2\pi\,\epsilonNu^{2}}{c\,(hc)^{3}}\,\Phi_{\IS,\ell}(\epsilonNu) \quad (\ell=0,1).  
\end{equation}
We tabulate the scattering opacities $\sigma_{\IS,0}$ and $\sigma_{\IS,1}$, which have units cm$^{-1}$, but use only $\sigma_{\IS,0}$ in the numerical experiments in Section~\ref{sec:numericalTableOpacities}, where $\sigma=\sigma_{\IS,0}$ in Equation~\eqref{eq:momentEquationMomentum}.

@@

The NES scattering kernel $R_{\NES}^{\Out}(\cos\alpha,\epsilonNu,\epsilonNu')$ is given by Eq.~(9) in \citet{mezzacappaBruenn_1993c}.  
As we did in the isoenergetic scattering case, we approximate the angular dependence of the scattering kernels by a two-term Legendre expansion; e.g.,
\begin{equation}
  R^{\Out}_{\NES}(\cos\alpha,\epsilonNu,\epsilonNu') 
  \approx \f{1}{2} \Phi_{\NES,0}^{\Out}(\epsilonNu,\epsilonNu') + \f{3}{2}\Phi^{\Out}_{\NES,1}(\epsilonNu,\epsilonNu')\,\cos\alpha,
\end{equation}
where
\begin{equation}
  \Phi_{\NES,\ell}^{\Out}(\epsilonNu,\epsilonNu')
  =\int_{-1}^{+1}R_{\NES}^{\Out}(\cos\alpha,\epsilonNu,\epsilonNu')\,P_{\ell}(\cos\alpha)\,d\cos\alpha \quad (\ell=0,1).  
\end{equation}
The in-scattering kernels $\Phi_{\NES,\ell}^{\In}(\epsilonNu,\epsilonNu')$ are obtained from symmetry properties \citep{cernohorsky_1994}
\begin{equation}
  \Phi_{\NES,\ell}^{\In}(\epsilonNu,\epsilonNu')=\Phi_{\NES,\ell}^{\Out}(\epsilonNu',\epsilonNu)
  \quad\text{and}\quad
  \Phi_{\NES,\ell}^{\Out}(\epsilonNu',\epsilonNu)
  =e^{-(\epsilonNu-\epsilonNu')/k_{\mbox{\tiny B}}T}\,\Phi_{\NES,\ell}^{\Out}(\epsilonNu,\epsilonNu').
  \label{eq:symmetryNES}
\end{equation}
Inserting the Legendre expansion into Eq.~\eqref{eq:homogeneousBoltzmann}, and taking the zeroth moment (number equation), gives
\begin{align}
  \f{1}{c}\deriv{\cJ}{t}
  &=\big(\,4\pi-\cJ(\epsilonNu)\,\big)\int_{\bbR^{+}}R_{\NES,0}^{\In}(\epsilonNu,\epsilonNu')\,\cJ(\epsilonNu')\,dV_{\epsilonNu'}
  -\bcH(\epsilonNu)\cdot\int_{\bbR^{+}}R_{\NES,1}^{\In}(\epsilonNu,\epsilonNu')\,\bcH(\epsilonNu')\,dV_{\epsilonNu'} \nonumber \\
  & \hspace{12pt}
  -\cJ(\epsilonNu)\int_{\bbR^{+}}R_{\NES,0}^{\Out}(\epsilonNu,\epsilonNu')\,\big(\,4\pi-\cJ(\epsilonNu')\,\big)\,dV_{\epsilonNu'}
  +\bcH(\epsilonNu)\cdot\int_{\bbR^{+}}R_{\NES,1}^{\Out}(\epsilonNu,\epsilonNu')\,\bcH(\epsilonNu')\,dV_{\epsilonNu'}.  
\end{align}
Similarly, taking the first moment (number flux) of Eq.~\eqref{eq:homogeneousBoltzmann} results in
\begin{align}
  \f{1}{c}\deriv{\bcH}{t}
  &=-\bcH(\epsilonNu)\int_{\bbR^{+}}R_{\NES,0}^{\In}(\epsilonNu,\epsilonNu')\,\cJ(\epsilonNu')\,dV_{\epsilonNu'}
  +\big(\,(4\pi/3)\,\bI-\bcK\,\big)\cdot\int_{\bbR^{+}}R_{\NES,1}^{\In}(\epsilonNu,\epsilonNu')\,\bcH(\epsilonNu')\,dV_{\epsilonNu'} \nonumber \\
  & \hspace{12pt}
  -\bcH\int_{\bbR^{+}}R_{\NES,0}^{\Out}(\epsilonNu,\epsilonNu')\,\big(\,4\pi-\cJ(\epsilonNu')\,\big)\,dV_{\epsilonNu'}
  +\bcK\cdot\int_{\bbR^{+}}R_{\NES,1}^{\Out}(\epsilonNu,\epsilonNu')\,\bcH(\epsilonNu')\,dV_{\epsilonNu'},
\end{align}
where we have defined
\begin{equation}
  R_{\NES,0}^{\In/\Out}
  =\f{1}{c\,(hc)^{3}}\,\f{1}{2}\,\Phi_{\NES,0}^{\In/\Out}(\epsilonNu,\epsilonNu')
  \quad\text{and}\quad
  R_{\NES,1}^{\In/\Out}
  =\f{1}{c\,(hc)^{3}}\,\f{3}{2}\,\Phi_{\NES,1}^{\In/\Out}(\epsilonNu,\epsilonNu').  
\end{equation}
We tabulate the scattering kernels $R_{\NES,0}^{\Out}$ and $R_{\NES,1}^{\Out}$, which have units MeV$^{-3}$ cm$^{-1}$, and obtain the in-scattering rates from the symmetry properties in Eq.~\eqref{eq:symmetryNES}.  
We only use $R_{\NES,0}^{\In/\Out}$ in the numerical experiments in Section~\ref{sec:numericalTableOpacities}, so that $R^{\In/\Out}=R_{\NES,0}^{\In/\Out}$ in Eqs~\eqref{eq:momentEquationEnergy} and \eqref{eq:momentEquationMomentum}.

>>>>

As we mentioned in Section~\ref{sec:twoMomentModel} when discussing Eqs.~\eqref{eq:momentEquationEnergy} and \eqref{eq:momentEquationMomentum}, the total opacity (excluding NES) is $\kappa(\epsilonNu,\bU)=\tilde{\chi}(\epsilonNu,\bU)+\sigma(\epsilonNu,\bU)$, where the absorption opacity $\tilde{\chi}(\epsilonNu,\bU)$, corrected for stimulated absorption, is given by the sum of the neutrino emissivity $\eta(\epsilonNu,\bU)$ and the neutrino absorptivity $\chi(\epsilonNu,\bU)$
\begin{equation}
  \tilde{\chi}(\epsilonNu,\bU) = \eta(\epsilonNu,\bU) + \chi(\epsilonNu,\bU). 
\end{equation}
The scattering opacity $\sigma(\epsilonNu,\bU)$ is due to (isotropic) elastic scattering.   
Neutrino-electron scattering gives rise to the integral operators involving the rates $R^{\In}(\epsilonNu,\epsilonNu',\bU)$ and $R^{\In}(\epsilonNu,\epsilonNu',\bU)$.  
(In the following, to simplify the notation, we will drop the explicit spatial dependence and dependence on material properties $\bU$.)  
To arrive at the collision terms for the two-moment model, we consider the space-homogeneous Boltzmann equation \citep[e.g., Eq.~(7) in][]{mezzacappaBruenn_1993c}
\begin{align}
  \f{1}{c}\deriv{f}{t}(\omegaNu,\epsilonNu,t)
  & = \eta(\epsilonNu) - \tilde{\chi}(\epsilonNu)\,f(\omegaNu,\epsilonNu,t) \nonumber \\
  & \hspace{12pt}
  + \frac{\epsilonNu^2}{c\,(hc)^3} \int_{\bbS^{2}} R_{\IS}(\bl'\cdot\bl,\epsilonNu)\,f(\omega', \epsilonNu,t)\,d\omegaNu'
  - \frac{\epsilonNu^2}{c\,(hc)^3} f(\omegaNu,\epsilonNu,t) \int_{\bbS^{2}} R_{\IS}(\bl'\cdot\bl,\epsilonNu)\,d\omega' \nonumber \\
  & \hspace{12pt} 
  + \frac{1}{c\,(hc)^3}\big(1- f(\omegaNu,\epsilonNu)\big) \, 
  \int_{\bbR^{+}} \int_{\bbS^{2}}  R^{\In}_{\NES} (\bl'\cdot\bl,\epsilonNu,\epsilonNu')\,f(\omega',\epsilonNu',t)\,d\epsilonNu' \epsilonNu'^{2}\,d\omega' \nonumber \\
  & \hspace{12pt} 
  - \frac{1}{c\,(hc)^3} f(\omegaNu,\epsilonNu,t) \int_{\bbR^{+}} 
  \int_{\bbS^{2}}  R^{\Out}_{\NES} (\bl'\cdot\bl,\epsilonNu,\epsilonNu^{\prime})\,\big(1-f(\omegaNu',\epsilonNu',t )\big)\,d\epsilonNu' \epsilonNu'^{2}\,d\omegaNu',
  \label{eq:homogeneousBoltzmann}
\end{align}
where we have brought back physical constants; i.e., the speed of light $c$ and the Planck constant $h$.  
In the scattering kernels, $\bl(\omegaNu)=(\cos\vartheta,\sin\vartheta\cos\varphi,\sin\vartheta\sin\varphi)^{\mbox{\tiny T}}$, so that
\begin{equation}
 \bl'\cdot\bl  = \mu\mu' + [(1-\mu^2)(1-\mu'^{2})]^{1/2}\cos(\phiNu-\phiNu')
 \equiv \cos\alpha,
 \quad(\mu=\cos\thetaNu)
\end{equation}
is the cosine of the angle between incoming and outgoing neutrino is the scattering process.  
In Eq.~\eqref{eq:homogeneousBoltzmann}, the first two terms are due to emission and absorption, the third and fourth terms are due to elastic (isoenergetic) neutrino scattering on nucleons and nuclei, and the last two terms are due to inelastic (non-isoenergetic) neutrino-electron scattering.

In the following subsections we separately give the collision terms for the two-moment model by taking angular moments of Eq.~\eqref{eq:homogeneousBoltzmann}.  
(The total collision term in the two-moment model is then given by the sum of the separate terms.)  
The opacities are computed using the SFHo equation of state \citep{steiner_etal_2013}, and tabulated for densities $\rho\in[1.66\times10^{4},3.16\times10^{15}]$~g~cm$^{-3}$, temperatures $T\in[1.16\times10^{9},1.84\times10^{12}]$~K, and electron fractions $Y_{e}\in[0.01,0.6]$.  
The neutrino energies are in the range $\epsilonNu\in[0.1,300]$~MeV.  
We use logarithmic grids to cover the range in density $N_{\rho}=308$, temperature ($N_{T}=129$), and the neutrino energy ($N_{\epsilonNu}=40$), and a linear grid to cover the range in electron fraction ($N_{Y_{e}}=60$).  
The neutrino electron scattering rates are tabulated in terms of neutrino energies ($\epsilonNu$ and $\epsilonNu'$), temperature, and $\eta=\mu_{\nu}/(k_{\mbox{\tiny B}}T)$.  
Here we use a logarithmic grid in $\eta\in[1.0\times10^{-8},2.5\times10^{3}]$~MeV, using $N_{\eta}=100$ points.

@@

Electron capture on nucleons and nuclei ($p+e^{-}\rightleftharpoons n+\nu_{e}$) is governed by the corrected absorption opacity $\tilde{\chi}(\epsilonNu)$ 
\begin{equation}
  \tilde{\chi}(\epsilonNu) = \chi(\epsilonNu) + \eta(\epsilonNu), 
\end{equation}
where $\eta(\epsilonNu)$ is the emissivity (taken as the sum of contributions from nucleons and nuclei)
\begin{equation}
  \eta = \eta_{\mbox{\tiny nucleon}} + \eta_{\mbox{\tiny nuclei}},
\end{equation}
and $ \chi(\epsilonNu)$ is the absorption opacity.  
They are related by equilibrium condition
\begin{equation}
  \chi(\epsilonNu) = e^{(\epsilonNu-\mu_{\nu})/k_{\mbox{\tiny B}}T}\,\eta(\epsilonNu). 
\end{equation}
Therefore,
\begin{equation}
  \chi = e^{(\epsilonNu-\mu_{\nu})/k_{\mbox{\tiny B}}T}\,(\eta_{\mbox{\tiny nucleon}} + \eta_{\mbox{\tiny nuclei}}) 
  \equiv \chi_{\mbox{\tiny nucleon}} + \chi_{\mbox{\tiny nuclei}},
\end{equation}
and
\begin{equation}
  \eta-\tilde{\chi}\,f=\tilde{\chi}\,\big(\,\eta/\tilde{\chi}-f\,\big)=\tilde{\chi}\,\big(\,f_{0}-f\,\big).  
\end{equation}
Thus, the corresponding terms in the moment equations are
\begin{equation}
  \f{1}{c}\deriv{\cJ}{t}=\tilde{\chi}\,\big(\,\cJ_{0}-\cJ\,\big)
  \quad\text{and}\quad
  \f{1}{c}\deriv{\bcH}{t}=-\tilde{\chi}\,\bcH.  
\end{equation}
We use the electron neutrino nucleon absorption opacity $\chi_{\mbox{\tiny nucleon}}$ and emissivity $\eta_{\mbox{\tiny nucleon}}$ as given by Eq.~(C13) and Eq.~(C15), respectively, in \citet{bruenn_1985}, and the electron-neutrino nuclei absorption opacity $\chi_{\mbox{\tiny nuclei}}$ and emissivity $\eta_{\mbox{\tiny nuclei}}$ as given by Eq.~(C29) and Eq.~(C27), respectively, in \citet{bruenn_1985}.  
We only tabulate the corrected absorption opacity $\tilde{\chi}$, which has units cm$^{-1}$.

@@

The isoenergetic scattering kernel $R_{\IS}(\bl'\cdot\bl,\epsilonNu)$ is taken as the sum of contributions from scattering on nucleons and nuclei
\begin{equation}
  R_{\IS}(\bl'\cdot\bl,\epsilonNu,\epsilonNu) = R_{\IS}^{\mbox{\tiny nucleon}}(\bl'\cdot\bl,\epsilonNu) + R_{\IS}^{\mbox{\tiny nuclei}}(\bl'\cdot\bl,\epsilonNu),
\end{equation}
where $R_{\IS}^{\mbox{\tiny nucleon}}$ and $R_{\IS}^{\mbox{\tiny nuclei}}$ are given by Eq. (35) and Eq. (36), respectively, in \citet{mezzacappaBruenn_1993a}.  
For inclusion in the two-moment model, these kernels are approximated by a two-term Legendre series
\begin{align}
  R_{\IS}(\cos\alpha,\epsilonNu) \approx \f{1}{2} \Phi_{\IS,0}(\epsilonNu) + \f{3}{2}\Phi_{\IS,1}(\epsilonNu)\,\cos\alpha.
\end{align}
By orthogonality of the Legendre polynomials ($P_{0}(x)=1,P_{1}(x)=x$), the expansion coefficients are given by
\begin{equation}
  \Phi_{\IS,\ell}(\epsilonNu)
  =\int_{-1}^{+1}R_{\IS}(\cos\alpha,\epsilonNu)\,P_{\ell}(\cos\alpha)\,d\cos\alpha \quad (\ell=0,1).  
\end{equation}
These integrals are performed numerically using a ??-point Gauss-Legendre quadrature rule.  (\ee{Ran: Fill in the number of quadrature points}.)
Inserting the Legendre expansion into Eq.~\eqref{eq:homogeneousBoltzmann} and taking the angular moments gives no contribution to the equation for $\cJ$ (energy conservation), while the contribution to the equation for $\bcH$ is
\begin{equation}
  \f{1}{c}\deriv{\bcH}{t}
  =-\big(\,\sigma_{\IS,0}-\sigma_{\IS,1}\big)\,\bcH,
\end{equation}
where we have defined
\begin{equation}
  \sigma_{\IS,\ell}(\epsilonNu)
  =\f{2\pi\,\epsilonNu^{2}}{c\,(hc)^{3}}\,\Phi_{\IS,\ell}(\epsilonNu) \quad (\ell=0,1).  
\end{equation}
We tabulate the scattering opacities $\sigma_{\IS,0}$ and $\sigma_{\IS,1}$, which have units cm$^{-1}$, but use only $\sigma_{\IS,0}$ in the numerical experiments in Section~\ref{sec:numericalTableOpacities}, where $\sigma=\sigma_{\IS,0}$ in Equation~\eqref{eq:momentEquationMomentum}.

@@

The NES scattering kernel $R_{\NES}^{\Out}(\cos\alpha,\epsilonNu,\epsilonNu')$ is given by Eq.~(9) in \citet{mezzacappaBruenn_1993c}.  
As we did in the isoenergetic scattering case, we approximate the angular dependence of the scattering kernels by a two-term Legendre expansion; e.g.,
\begin{equation}
  R^{\Out}_{\NES}(\cos\alpha,\epsilonNu,\epsilonNu') 
  \approx \f{1}{2} \Phi_{\NES,0}^{\Out}(\epsilonNu,\epsilonNu') + \f{3}{2}\Phi^{\Out}_{\NES,1}(\epsilonNu,\epsilonNu')\,\cos\alpha,
\end{equation}
where
\begin{equation}
  \Phi_{\NES,\ell}^{\Out}(\epsilonNu,\epsilonNu')
  =\int_{-1}^{+1}R_{\NES}^{\Out}(\cos\alpha,\epsilonNu,\epsilonNu')\,P_{\ell}(\cos\alpha)\,d\cos\alpha \quad (\ell=0,1).  
\end{equation}
(Again, these integrals are computed numerically using a ??-point Gauss-Legendre quadrature.  (\ee{Ran: Fill in the number of quadrature points}.))
The in-scattering kernels $\Phi_{\NES,\ell}^{\In}(\epsilonNu,\epsilonNu')$ are obtained from symmetry properties \citep{cernohorsky_1994}
\begin{equation}
  \Phi_{\NES,\ell}^{\In}(\epsilonNu,\epsilonNu')=\Phi_{\NES,\ell}^{\Out}(\epsilonNu',\epsilonNu)
  \quad\text{and}\quad
  \Phi_{\NES,\ell}^{\Out}(\epsilonNu',\epsilonNu)
  =e^{-(\epsilonNu-\epsilonNu')/k_{\mbox{\tiny B}}T}\,\Phi_{\NES,\ell}^{\Out}(\epsilonNu,\epsilonNu').
  \label{eq:symmetryNES}
\end{equation}
Inserting the Legendre expansion into Eq.~\eqref{eq:homogeneousBoltzmann}, and taking the zeroth moment (number equation), gives
\begin{align}
  \f{1}{c}\deriv{\cJ}{t}
  &=\big(\,4\pi-\cJ(\epsilonNu)\,\big)\int_{\bbR^{+}}R_{\NES,0}^{\In}(\epsilonNu,\epsilonNu')\,\cJ(\epsilonNu')\,dV_{\epsilonNu'}
  -\bcH(\epsilonNu)\cdot\int_{\bbR^{+}}R_{\NES,1}^{\In}(\epsilonNu,\epsilonNu')\,\bcH(\epsilonNu')\,dV_{\epsilonNu'} \nonumber \\
  & \hspace{12pt}
  -\cJ(\epsilonNu)\int_{\bbR^{+}}R_{\NES,0}^{\Out}(\epsilonNu,\epsilonNu')\,\big(\,4\pi-\cJ(\epsilonNu')\,\big)\,dV_{\epsilonNu'}
  +\bcH(\epsilonNu)\cdot\int_{\bbR^{+}}R_{\NES,1}^{\Out}(\epsilonNu,\epsilonNu')\,\bcH(\epsilonNu')\,dV_{\epsilonNu'}.  
\end{align}
Similarly, taking the first moment (number flux) of Eq.~\eqref{eq:homogeneousBoltzmann} results in
\begin{align}
  \f{1}{c}\deriv{\bcH}{t}
  &=-\bcH(\epsilonNu)\int_{\bbR^{+}}R_{\NES,0}^{\In}(\epsilonNu,\epsilonNu')\,\cJ(\epsilonNu')\,dV_{\epsilonNu'}
  +\big(\,(4\pi/3)\,\bI-\bcK\,\big)\cdot\int_{\bbR^{+}}R_{\NES,1}^{\In}(\epsilonNu,\epsilonNu')\,\bcH(\epsilonNu')\,dV_{\epsilonNu'} \nonumber \\
  & \hspace{12pt}
  -\bcH\int_{\bbR^{+}}R_{\NES,0}^{\Out}(\epsilonNu,\epsilonNu')\,\big(\,4\pi-\cJ(\epsilonNu')\,\big)\,dV_{\epsilonNu'}
  +\bcK\cdot\int_{\bbR^{+}}R_{\NES,1}^{\Out}(\epsilonNu,\epsilonNu')\,\bcH(\epsilonNu')\,dV_{\epsilonNu'},
\end{align}
where we have defined
\begin{equation}
  R_{\NES,0}^{\In/\Out}
  =\f{1}{c\,(hc)^{3}}\,\f{1}{2}\,\Phi_{\NES,0}^{\In/\Out}(\epsilonNu,\epsilonNu')
  \quad\text{and}\quad
  R_{\NES,1}^{\In/\Out}
  =\f{1}{c\,(hc)^{3}}\,\f{3}{2}\,\Phi_{\NES,1}^{\In/\Out}(\epsilonNu,\epsilonNu').  
\end{equation}
We tabulate the scattering kernels $R_{\NES,0}^{\Out}$ and $R_{\NES,1}^{\Out}$, which have units MeV$^{-3}$ cm$^{-1}$, and obtain the in-scattering rates from the symmetry properties in Eq.~\eqref{eq:symmetryNES}.  
We only use $R_{\NES,0}^{\In/\Out}$ in the numerical experiments in Section~\ref{sec:numericalTableOpacities}, so that $R^{\In/\Out}=R_{\NES,0}^{\In/\Out}$ in Eqs~\eqref{eq:momentEquationEnergy} and \eqref{eq:momentEquationMomentum}.
